\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-05 pldi/draft.tex
Fix
Changes in the sort --parallel subsection

We attribute the small drop above $96\times$ in \ttt{spell} \ttt{sort} \ttt{wf}
and \ttt{bi-grams} to a combination of hyperthreading (128 virtual vs. 64
physical CPUs) \todo{the fixed width on \ttt{sort}}.
% I say "we attribute" because it's a way to say something without being certain
% FIXME: We need to take linear benchmarks on deathstar and livestar to see
% where it starts dropping
% not explain scope the results appropriately
% rinard

@@

@@

\heading{Parallel Sort}
We compare a single \todo{GNU} \ttt{sort} optimized by \sys with the
same \ttt{sort} with the \ttt{--parallel} flag set. The results are
shown in \Cref{fig:sort-parallel-comparison}. There are a couple
important points to note. First of all, \sys without the \ttt{eager}
optimization enabled performs comparably with \ttt{sort
  --parallel}. On the other hand, enabling the \ttt{eager}
optimization allows \sys to perform much better (almost \todo{2}
times) than \ttt{sort} with the \ttt{--parallel} flag. That is because
\sys with the \ttt{eager} optimization enabled essentially uses
intermediate buffers (on disk) between the merge phases. On the other
hand, \ttt{sort --parallel} sorts the input trying to minimize
additional memory (so doesn't allocate intermediate buffers). \kk{If
  we want, we can add here numbers where we allow sort --parallel to
  use additional ram for intermediate buffers (using the -S flag).} Of
course, it would be possible to manually create a longer pipeline and
use intermediate buffers to speed this up, but this defeats the
purpose of having a simple solution.

Finally, based on the comparison with \ttt{sort --parallel} it seems
that the scaleup limit that \sys reaches with sort is \todo{inherent
  and cannot be overpassed} \kk{We have to find a way to argue for the
  previous statement}. Because of that, all pipelines and scripts that
contain \ttt{sort} in our evaluation (several one-liners and unix50)
have this upper bound on scalability. \kk{Maybe move this paragraph in
  the one-liners section.}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
  \caption{ \textbf{Comparison with \ttt{sort --parallel}.}
  }
  \vspace{-15pt}
  \label{fig:sort-parallel-comparison}
\end{figure}

\kk{There is a weird thing in the results. That the sequential one
  with --parallel=1 takes more than the \sys sequential one.}

>>>>

We attribute the small drop above $96\times$ in \ttt{spell} \ttt{sort} \ttt{wf}
and \ttt{bi-grams} to a combination of hyperthreading (128 virtual vs. 64
physical CPUs) \todo{the fixed width on} \ttt{sort}.
% I say "we attribute" because it's a way to say something without being certain
% FIXME: We need to take linear benchmarks on deathstar and livestar to see
% where it starts dropping
% not explain scope the results appropriately
% rinard

@@

\kk{I think the scoping of this should change to be both
  micro-benchmarks and comparison with existing work. Especially since
  we will have some form of comparison with Raftlib and phoenix. I
  guess this section is trying to establish \sys limits by comparing
  it with similar (but not identical) work.}

@@

\heading{Parallel Sort}
%% \kk{Explain what is the purpose of the
%%   microbenchmark. We want to establish that \sys scalability is as
%%   good as it could be.}
We first compare a single GNU \ttt{sort} optimized by \sys with the
same \ttt{sort} with the \ttt{-{}-parallel} flag set. Even though the
\ttt{-{}-parallel} flag is not a general parallelism solution, this
comparison serves to establish a baseline for \sys's performance. The
results are shown in \Cref{fig:sort-parallel-comparison}. Note that
for \ttt{sort -{}-parallel} parallelism is set to double the number
set of \sys (i.e. the final point is for \ttt{-{}-parallelism=128})
since \sys also spawns separate processes for sorting and for
merging. There are a couple important points to note.

First of all, \sys without the \ttt{eager} optimization enabled
performs comparably with \ttt{sort -{}-parallel}. On the other hand,
enabling the \ttt{eager} optimization allows \sys to perform much
better ($\sim 2$ times) than \ttt{sort} with the \ttt{-{}-parallel}
flag. That is because \sys with the \ttt{eager} optimization enabled
essentially uses intermediate buffers (on disk) between the merge
phases.
%% On the other hand, \ttt{sort --parallel} sorts the input trying to
%% minimize additional memory (so doesn't allocate intermediate
%% buffers).
%% %% \kk{If we want, we can add here numbers where we allow sort --parallel
%% %%   to use additional ram for intermediate buffers (using the -S flag).}
%% Of course, it would be possible to manually create a longer pipeline
%% and use intermediate buffers to speed this up, but this defeats the
%% purpose of having a simple solution.

Second, based on the comparison with \ttt{sort --parallel} it seems
that the scaleup limit that \sys reaches with sort is \todo{inherent
  and cannot be overpassed} \kk{We have to find a way to argue for the
  previous statement}. Because of that, all pipelines and scripts that
contain \ttt{sort} in our evaluation (several one-liners and unix50)
have this upper bound on scalability. \kk{Maybe move this paragraph in
  the one-liners section.}

Finally, this comparison showcases the benefit that \sys provides to
command developers. Instead of implementing a custom flag that enables
parallelism for a command, a simple specification of the command
parallelizability allows \sys to scale it similarly to a handcrafted
implementation with almost no effort.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
  \caption{ \textbf{Comparison with \ttt{sort -{}-parallel}.}
  }
  \vspace{-15pt}
  \label{fig:sort-parallel-comparison}
\end{figure}

%% \kk{There is a weird thing in the results. That the sequential one
%%   with --parallel=1 takes more than the \sys sequential one.}
